---
title: "Introduction to Keywords"
date: "2023-09-04"
categories: ["Digital Advertising", "Keywords"]
toc: true
draft: true
---

These notes are based on google ads and my own reasearch. This is just an introduction post. Mostly notes. 

```{r keyword example table build}
#| echo: false
library(gt)
keyword_ex <- data.frame(
  type = c("Broadmatch",
           "Broadmath Modifiers +",
           "Phrase Match",
           "Exactmatch [ ]"),
  keyword = c("children's bicycles",
              "+children's +bicycles",
              "'children's bicycles'",
              "[children's bicycles]"),
  search_1 = c("buy kids bicycles",
               "children's toys and bicylces",
               "safe children's bicycles",
               "children's bicycles"),
  search_2 = c("children's scooters",
               "mountain bicycles for children",
               "buy children's bicycles",
               "kids bicycles"),
  search_3 = c("",
               "",
               "children's bicycles for sale",
               "bicycles for children")
)
keyword_ex_tb <- keyword_ex |> gt() |>
  tab_style(
    style = list(
      cell_fill(color = "white"),
      cell_text(color = "blue")
      ),
    locations = cells_body(rows = 1)
  ) |>
  tab_style(
    style = list(
      cell_fill(color = "white"),
      cell_text(color = "darkgreen")
      ),
    locations = cells_body(rows = 2)
  ) |>
  tab_style(
    style = list(
      cell_fill(color = "white"),
      cell_text(color = "purple")
      ),
    locations = cells_body(rows = 3)
  ) |>
  tab_style(
    style = list(
      cell_fill(color = "white"),
      cell_text(color = "orange")
      ),
    locations = cells_body(rows = 4)
  ) |>
  cols_label(
    type = "Keyword Type",
    keyword = "Keyword",
    search_1 = "Search 1",
    search_2 = "Search 2",
    search_3 = "Search 3"
  )
```


# Keywords

## Definitions

$\color{blue}\textbf{Broadmatch}$: Show ad if keyword (or any variation like misspellings, synonyms, or related search) are included in a user's search terms. 

$\color{green}\textbf{Broadmath Modifiers +}$: Show ad if keyword or close variations are apart of the search. 

$\color{purple}\textbf{Phrase Match " "}$: Shows ads if the keywords within the quotation marks or close variations of them match a user's search term. 

$\color{orange}\textbf{Exactmatch [ ]}$: Shows ads only if search means the same thing as your keyword. 

$\color{red}\textbf{Negative Keywords -}$: Ads wont show if someone includes that negative keyword in their search. 

## Search Examples

```{r keyword example table print}
#| echo: false
keyword_ex_tb
```

## Examples and Benefits

::: {.panel-tabset}

### X 

Click on each tab for examples and benefits. 

### $\color{blue}\textbf{Broadmatch}$

**Example**

* bike $\rightarrow$ bicycle 

* crane $\rightarrow$ lifting gear

**Benefits**

* Less time building keyword lists.

* Cast a wide net.

* Discover terms you might not be aware of. 

### $\color{green}\textbf{Broadmath Modifiers +}$

**Example**

* +bike, +girl, +bike $\rightarrow$ blue bicycle for 6 year old girls

* +bike, +girl, +bike $\nrightarrow$ girl bike 

* +john, +deere, +tractor $\rightarrow$ john deere tractor for home property

* +john, +deere, +tractor $\nrightarrow$ home property tractor

**Benefits**

* Discover terms you not be aware of. 

* Indicate if certain concepts must be present. 


### $\color{purple}\textbf{Phrase Match " "}$

**Example**

* "My First Big Bike" $\rightarrow$ My First Big Bike

* "My First Big Bike" $\nrightarrow$ first big bike 

* "The Weekender" $\rightarrow$ The Weekender

* "The Weekender" $\nrightarrow$ The Bar And Grill Weekender

**Benefits**

* No extra words can be added to phrase. 

* More targeted than broad match. 

* More flexible than exact match. 

:::

::: {.panel-tabset}

### X

Click on each tab for examples and benefits. 

### $\color{orange}\textbf{Exactmatch [ ]}$

**Example**

* [men's bicycles] $\rightarrow$ men's bicycles or men's bike

* [men's bicycles] $\nrightarrow$ men's bikes for sale

* [compact utility tractor] $\rightarrow$ compact utility tractor

* [compact utility tractor] $\nrightarrow$ compact utility tractor for lease 

**Benefits**

* Match with very specific searches. 

### $\color{red}\textbf{Negative Keywords -}$

**Example**

* (-scooter) $\nrightarrow$ scooter

* (-toy) $\nrightarrow$ toy

**Benefits**

* Match terms you know your customers will be using. 

:::

